A highly-available move operation for replicated trees

### 2019-06-21 paper/move-op.tex
Updated version of distributed filesystems.

Unison is a file synchronisation tool with a formal specification~\cite{PierceVouillon:UnisonSpecTR} which allows two replicas, such as laptops, desktops or servers, to synchronise the state of a directory. 
Unison permits offline updates to both replicas. 
The (possibly divergent) state of two replicas are resynchronised by running the Unison tool on the two replicas. 
Unison works by recording a summary record of the state of the filesystem on each replica,  inspecting this summary whenever the unison program is invoked, and sharing any update information with the other replica in order to determine the changes made to the filesystems on each replica since they were last synchronised. 
This approach has the benefit that no special kernel extension or daemon process is required; instead Unison is an application program which inspects the current state of the filesystem whenever it is run. 
When presented with the move operations described in Figure~\ref{fig:move-same-item}, Unison duplicates the files, resulting in the outcome shown in Figure~\ref{fig:move-same-item}~(a). 
Unison is unable to automatically synchronise the move operations shown in Figure~\ref{fig:move-cycle} and instead asks the user to confirm whether changes made to directories A and B should be propagated or deleted. 
As a result, depending on user choice, both replicas can end up in one of four states, including those shown in Figure~\ref{fig:move-cycle}~(b), (c) or (d) or with an empty root directory on both replicas. 
In all cases, Unison successfully ensures both replicas are in the same state after synchronisation.

Coda is a distributed client-server file system which allows clients to cache or \emph{hoard} copies of files otherwise stored in a shared data repository~\cite{kistler1992coda}. 
Clients can edit copies of data held in the cache while offline, during which time a kernel module keeps track of all updates. 
When the client comes back online, the client attempts to resynchronise changes with the repository. 
This design means that conflicts may arise if multiple clients edit the same files or directories while one or more clients are offline. 
One proposed solution to resolving conflicts in files is to support application-specific resolvers which ``encapsulate the knowledge needed for file resolution''~\cite{Kumar:1995wf}, an approach is similar to that found in Bayou~\cite{Terry:1995dn}.
Concurrent edits on directory entries across different replicas in Coda have been considered, but the authors note that they do ``not address transparent resolution of cross-directory [move operations] in [their] current implementation.''~\cite{kumar1993log}.
Furthermore, while the authors consider a number of conflicts associated with directory move operations, they do not highlight the potential for the creation of cycles as seen in Figure~\ref{fig:move-cycle} (a).

The Ficus replicated file system~\cite{Reiher:1994wh} is a in-kernel SunOS-based replicated peer-to-peer file system. 
Ficus supported updates to replicas during periods of network partition and claims ``conflicting updates to directories are detected an automatically repaired''~\cite{guy1990implementation}. 
Unfortunately we were unable to find a precise definition of the algorithm used in any of the available publications. 
Rumor~\cite{Guy:1999gy} is the successor to Ficus. 
An important distinction is between Rumor and previous work is that it operates at the application layer and needs to be invoked periodically by the user or a deamon process.
Rumor is therefore very similar in operation and design to Unison described earlier.

%The user manual (in Archive) is quite helpful for Rumor: https://web.archive.org/web/20170705140958/ftp://ftp.cs.ucla.edu/pub/rumor/rumor_users_manual.ps. Note that this also makes the claim that "Unix directories are an important example of a kind of file for which Rumor automatically resolves conflicts"

TODO: Roam~\cite{Ratner:1999fh}

@@

TODO

>>>>

Coda is a distributed client-server file system which allows clients to cache or \emph{hoard} copies of files otherwise stored in a shared data repository~\cite{kistler1992coda}. 
Clients can edit copies of data held in the cache while offline, during which time a kernel module keeps track of all updates. 
When the client comes back online, the client attempts to resynchronise changes with the repository. 
This design means that conflicts may arise if multiple clients edit the same files or directories while one or more clients are offline. 
One proposed solution to resolving conflicts in files is to support application-specific resolvers which ``encapsulate the knowledge needed for file resolution''~\cite{Kumar:1995wf}, an approach similar to that found in Bayou~\cite{Terry:1995dn}.
Concurrent edits on directory entries across different replicas in Coda have been considered, but the authors note that they do ``not address transparent resolution of cross-directory [move operations] in [their] current implementation''~\cite{kumar1993log}.
Furthermore, while the authors consider a number of conflicts associated with directory move operations, they do not highlight the potential for the creation of cycles as seen in Figure~\ref{fig:move-cycle} (a).

The Ficus replicated file system~\cite{Reiher:1994wh} was a in-kernel SunOS-based replicated peer-to-peer file system. 
Ficus supported updates to replicas during periods of network partition and claims ``conflicting updates to directories are detected and automatically repaired''~\cite{guy1990implementation}. 
Unfortunately we were unable to find a precise definition of the algorithm used in any of the available publications. 
Rumor~\cite{Guy:1999gy} is the successor to Ficus. 
An important distinction between Rumor and previous work is that it operates at the application layer and needs to be invoked periodically by the user or by a deamon process. 
The original version of Rumor was unable to scale beyond 20 replicas, however an extension called Roam~\cite{Ratner:1999fh} allowed an order of magnitude better scaling. 
We obtained the user manual and source code for Rumor from \url{archive.org}, however after modest effort, we were unable to get the software running.
%The user manual (in Archive) is quite helpful for Rumor: https://web.archive.org/web/20170705140958/ftp://ftp.cs.ucla.edu/pub/rumor/rumor_users_manual.ps. Note that this also makes the claim that "Unix directories are an important example of a kind of file for which Rumor automatically resolves conflicts"

Unison is a file synchronisation tool with a formal specification~\cite{PierceVouillon:UnisonSpecTR} which allows two replicas, such as laptops, desktops or servers, to synchronise the state of a directory.
Unison permits offline updates to both replicas.
Like Rumor, Unison also works at the application layer and supports the resynchronisation of two (possibly divergent) replicas by executing a copy of Unison on both nodes. 
Unison records a summary record of the state of the filesystem on each replica whenever it is run.
By inspecting this summary whenever the Unison program is invoked, and by sharing any update information with the other replica, both replicas can determine the changes made to the filesystems on each replica since they were last synchronised. 
This approach has the benefit that no special kernel extension or daemon process is required; instead the user invokes the Unison program whenever synchronisation is needed.
When presented with the move operations described in Figure~\ref{fig:move-same-item}, Unison duplicates the files, resulting in the outcome shown in Figure~\ref{fig:move-same-item}~(a). 
Unison is unable to automatically synchronise the move operations shown in Figure~\ref{fig:move-cycle} and instead asks the user to confirm whether changes made to directories A and B should be propagated or deleted. 
As a result, depending on user choice, both replicas can end up in one of four states, including those shown in Figure~\ref{fig:move-cycle}~(b), (c) or (d) or with an empty root directory on both replicas. 
In all cases, Unison successfully ensures both replicas are in the same state after synchronisation.

@@

TODO.
